---
title: "arg_check(): more examples"
---
<br>

```{r, include = FALSE}
knitr::opts_chunk$set(
    error = TRUE, 
    collapse = FALSE,
    comment = "##"
)
```

```{r setup}
library(saferDev)
```
<br>


### Dataset

```{r}
vec1 <- -1:3 # vector of integers.
vec2 <- 1:3 / 3 # vector of positive proportions.
vec3 <- c(1, 2, 3) # vector of positive integers but stored as "double".
vec4 <- "pearson" # vector of characters.
vec5 <- c("a", "b","a", "b") # vector of characters.
mat1 <- matrix(vec1) # matrix of integers.
mat2 <- matrix(c(1:3 / 3, NA)) # matrix of proportions with NA.
```
<br>

### Dataset info

```{r}
vec1 # vector of integers.
vec2 # vector of proportions.
vec3 # vector of integers but stored as "double".
vec4 # vector of characters.
vec5 # vector of characters.
mat1 # matrix of integers.
mat2 # matrix of proportions with NA.
```
<br>

### Simple examples
```{r}
# Check that vec1 is type integer.
arg_check(data = vec1, typeof = "integer")


# Check that vec1 is a vector of type double -> error because vec1 is not type "double", 
# but type "integer".
arg_check(data = vec1, class = "vector", typeof = "double")


# Check that vec1 is an integer vector of length 3 without negative values and 
# without NA -> error because of length 5 and negative values inside vec1.
arg_check(
    data = vec1, 
    class = "vector", 
    typeof = "integer", 
    length = 3, 
    neg_values = FALSE, 
    na_contain = FALSE
)


# No result displayed because the output list is assigned into res
# (see below the print argument).
res <- arg_check(data = vec1, typeof = "integer")


# With data = NULL, the function systematically report a checking problem.
arg_check(data = NULL, typeof = "integer")
```
<br>

### Arguments <code>class</code>, <code>typeof</code>, <code>mode</code> and <code>length</code>
These arguments accept as values the classical kind of class, type, mode and length, respectively, except the <code>class</code> argument, which:

- also accepts <code>"vector"</code>, <code>"ggplot2"</code> (i.e., objects of class <code>c("gg", "ggplot")</code> for R < 4.5 and <code>c("ggplot2::ggplot", "ggplot", "ggplot2::gg", "S7_object", "gg")</code> otherwise) or <code>"ggplot_built"</code> (i.e., objects of class <code>"ggplot_built"</code> for R < 4.5 and <code>c("ggplot2::ggplot_built", "ggplot_built", "ggplot2::gg", "S7_object")</code> otherwise) .
- accepts <code>"matrix"</code> for matrices (objects of class <code>c("matrix" "array")</code>).


```{r, echo = TRUE, error = TRUE}
# Example.
arg_check(
    data = vec1, 
    class = "vector", 
    typeof = "integer", 
    mode = "numeric", 
    length = 5, 
)

# Warning: the function does not check for inconsistencies between arguments.
# It just checks if everything is ok between arguments values and the tested object.
# Here the inconsistancy between typeof and mode is not signaled.
arg_check(
    data = vec1, 
    typeof = "integer", 
    mode = "character"
)

# Bad mode argument value.
arg_check(
    data = vec1, 
    mode = "integer",
)
```
<br>

### Argument <code>prop</code>
```{r}
# Check for values between 0 and 1 only.
arg_check(
    data = vec2, 
    prop = TRUE 
)
```
<br>

### Argument <code>double_as_integer_allowed</code>
```{r}
# Check for integers, even if numbers are stored as double.
# With double_as_integer_allowed = TRUE # with TRUE, integers stored as double are accepted.
arg_check(
    data = vec3, 
    typeof = "integer",
    double_as_integer_allowed = TRUE
)
```
<br>

### Argument <code>options</code>
```{r}
# Check for the content among values. Here no error detected because vec4 contains "pearson".
arg_check(
    data = vec4, 
    options = c("pearson", "spearman", "kendall")
)
```
<br>

### Argument <code>all_options_in_data</code>
```{r}
# Check for the content among values.
# Here no error detected because vec5 indeed contains at least one copy of "a" and "b" and nothing else.
arg_check(
    data = vec5,
    options = c("a", "b"), 
    all_options_in_data = TRUE
)

# Here no error detected because vec5 indeed contains at least one copy of "a" and "b", among other values.
arg_check(
    data = vec5,
    options = c("a", "b", "c"), 
    all_options_in_data = FALSE
)

# Here an error is detected because vec5 must contain "a", "b" and "c", at least once each and nothing else.
arg_check(
    data = vec5,
    options = c("a", "b", "c"), 
    all_options_in_data = TRUE
)
```
<br>

### Argument <code>na_contain</code>
```{r}
# Check for the absence of NA.
# Warning: na_contain = TRUE means "no check". Thus, NA can be present or not. 
arg_check(
    data = mat2, 
    class = "matrix", 
    prop = TRUE,
    na_contain = FALSE 
)
```
<br>

### Argument <code>neg_values</code>
```{r}
# Check for the absence of negative values.
# Warning: neg_values = TRUE means "no check". Thus, negative numbers can be present or not. 
arg_check(
    data = vec1, 
    typeof = "integer",
    neg_values = FALSE
)

# Warning: neg_values = FALSE cannot be used if typeof and mode are NULL. 
arg_check(
    data = mat1, 
    class = "matrix",
    neg_values = FALSE
)
```
<br>

### Argument <code>inf_values</code>
```{r}
# Check for the absence of Inf/-Inf values.
# Warning: inf_values = TRUE means "no check". Thus, infinite numbers can be present or not.
arg_check(
    data = vec2, 
    typeof = "double",
    inf_values = FALSE
)

# Warning: inf_values = FALSE cannot be used with typeof = "integer". 
arg_check(
    data = mat1, 
    typeof = "integer",
    inf_values = FALSE
)

# Warning: inf_values = FALSE cannot be used if typeof and mode are NULL. 
arg_check(
    data = mat1, 
    class = "matrix",
    inf_values = FALSE
)
```
<br>

### Argument <code>print</code>
```{r}
# No error message printed because print is FALSE.
res <- arg_check(
    data = mat1, 
    class = "data.frame",
    print = FALSE
)

# Error message printed. Warning: does not trigger "exit". It is just a message.
res <- arg_check(
    data = mat1, 
    class = "data.frame",
    print = TRUE
)

# Even if print is TRUE, no error message printed because no error.
res <- arg_check(
    data = mat1, 
    class = "matrix",
    print = TRUE
)
```
<br>

### Argument <code>data_name</code>
```{r}
# Modification of the name of the tested object in the message
arg_check(
    data = vec1, 
    typeof = "integer", 
    data_name = "OBSERVATION_1"
)

# This argument is interesting when data is not an object.
x <- 1
arg_check(
    data = get("x"), 
    typeof = "integer", 
    data_name = "x"
)

# If "data_name" is not used.
arg_check(
    data = get("x"), 
    typeof = "integer", 
    data_name = NULL
)
```
<br>

### Arguments <code>data_arg</code>
```{r}
# Message with "argument". Useful when arg_check() is used inside a function to test its arguments.
arg_check(
    data = vec2, 
    typeof = "integer", 
    data_arg = TRUE
)

# Message with "object".
arg_check(
    data = vec2, 
    typeof = "integer", 
    data_arg = FALSE
)
```
<br>

### Argument <code>safer_check</code>
```{r}
# Safer checkings performed before main code running.
arg_check(
    data = mat1, 
    class = "matrix", 
    safer_check = TRUE
)
```
<br>

### Argument <code>lib_path</code>
```{r}
# The "NOTGOOD" path does not exists.
arg_check(
    data = mat1, 
    class = "matrix", 
    lib_path = "NOTGOOD"
)
```
<br>

### Argument <code>error_text</code>
```{r}
# Add a text in error messages returned by all_args_here().
arg_check(
    data = mat1, 
    class = "matrix", 
    lib_path = "NOTGOOD",
    error_text = " === TEXT ADDED ==="
)
```
<br>

### All the arguments
```{r}
arg_check(
    data = vec1, # object to check.
    class = "vector", # check the expected class.
    typeof = "integer", # check the expected type.
    mode = NULL,  # check the expected mode.
    length = NULL,  # check the expected length.
    prop = FALSE,  # check the expected proportion.
    double_as_integer_allowed = FALSE,  # can integers be typeof = "double" or class = "numeric"?
    options = NULL,  # check all the possible option values.
    all_options_in_data = FALSE,  # all the options argument values must be present at least once in the data argument, and nothing else?
    na_contain = FALSE,  # can data contain NA? Only tested if FALSE.
    neg_values = TRUE,  # can data contain negative values? Only tested if FALSE.
    inf_values = TRUE,  # can data contain infinite values? Only tested if FALSE.
    print = FALSE,  # print a message if the $problem output is TRUE? 
    data_name = NULL,  # name of the object to test.
    data_arg = TRUE,  # If TRUE, "ARGUMENT" is written in output messages, otherwise "OBJECT".
    safer_check = TRUE, # perform some "safer" checks (see https://github.com/safer-r)?
    lib_path = NULL, # where the required packages are if not in the default directories.
    error_text = "" # add information in error messages returned by the function.
)
```
