---
title: "arg_check(): more examples"
---
<br>

```{r, include = FALSE}
knitr::opts_chunk$set(
    error = TRUE, 
    collapse = FALSE,
    comment = "##"
)
```

```{r setup}
library(saferDev)
```
<br>


### Dataset

```{r}
vec1 <- -1:3 # vector of integers.
vec2 <- 1:3 / 3 # vector of proportions.
vec3 <- c(1, 2, 3) # vector of integers but stored as "double".
vec4 <- "pearson" # vector of characters.
vec5 <- c("a", "b","a", "b") # vector of characters.
mat1 <- matrix(vec1) # matrix of integers.
mat2 <- matrix(c(1:3 / 3, NA)) # matrix of proportions with NA.
```
<br>

### Dataset info

```{r}
vec1 # vector of integers.
vec2 # vector of proportions.
vec3 # vector of integers but stored as "double".
vec4 # vector of characters.
vec5 # vector of characters.
mat1 # matrix of integers.
mat2 # matrix of proportions with NA.
```
<br>

### Simple examples
```{r}
# Check that vec1 is of class integer.
arg_check(data = vec1, class = "integer")


# Check that vec1 is a numeric vector -> error because vec1 is a vector of integers.
arg_check(data = vec1, class = "numeric")


# Check that vec1 is an integer vector of length 3 without negative values and without NA -> error because of length 5 and negative values inside vec1.
arg_check(data = vec1, class = "vector", typeof = "integer", length = 3, neg_values = FALSE, na_contain = FALSE)


# No result displayed because the output list is assigned into res (see below the print argument).
res <- arg_check(data = vec1, class = "integer")


# With data = NULL, the function systematically report a checking problem.
arg_check(data = NULL, class = "integer")
```
<br>

### Arguments <code>class</code>, <code>typeof</code>, <code>mode</code> and <code>length</code>
As value, they accept the classical kind of class, type, mode and legnth, respectively, except the <code>class</code> argument, which: 
\itemize{
    \item has also accepts \code{"vector"}, \code{"ggplot2"} (i.e., objects of class \code{c("gg", "ggplot")} for R < 4.5 and \code{c("ggplot2::ggplot", "ggplot", "ggplot2::gg", "S7_object", "gg")} otherwise) or \code{"ggplot_built"} (i.e., objects of class \code{"ggplot_built"} for R < 4.5 and \code{c("ggplot2::ggplot_built", "ggplot_built", "ggplot2::gg", "S7_object")} otherwise) 
    \item accepts "matrix" for matrices (and not "matrix" "array").
}

```{r, echo = TRUE, error = TRUE}
# Example.
arg_check(
    data = vec1, 
    class = "vector", 
    typeof = "integer", 
    mode = "numeric", 
    length = 5, 
)

# Warning: the function does not check for inconsistencies between arguments. It just checks if everything is ok between arguments values and the tested object.
arg_check(
    data = vec1, 
    typeof = "integer", 
    mode = "character", # the mode "character" exists but is inconsistant with typeof "integer". However, this aspect is not signaled by the arg_check().
)

# Bad mode argument value.
arg_check(
    data = vec1, 
    mode = "integer",
)
```
<br>

### Argument <code>prop</code>
```{r}
# Check for values between 0 and 1 only.
arg_check(
    data = vec2, 
    prop = TRUE 
)
```
<br>

### Argument <code>double_as_integer_allowed</code>
```{r}
# Check for integers, even if numbers are stored as double.
arg_check(
    data = vec3, 
    typeof = "integer",
    double_as_integer_allowed = TRUE # with TRUE, integers stored as double are accepted
)
```
<br>

### Argument <code>options</code>
```{r}
# Check for the content among values.
arg_check(
    data = vec4, 
    options = c("pearson", "spearman", "kendall")
)
```
<br>

### Argument <code>all_options_in_data</code>
```{r}
# Check for the presence of values.
arg_check(
    data = vec5,
    options = c("a", "b"), 
    all_options_in_data = TRUE
)

# Check for the content among values.
arg_check(
    data = vec5,
    options = c("a", "b", "c"), 
    all_options_in_data = FALSE
)

# Check for the presence of values.
arg_check(
    data = vec5,
    options = c("a", "b", "c"), 
    all_options_in_data = TRUE
)
```
<br>

### Argument <code>na_contain</code>
```{r}
# Check for the absence of NA. Warning: na_contain = TRUE means "no check". Thus, NA can be present or not. 
arg_check(
    data = mat2, 
    class = "matrix", 
    prop = TRUE,
    na_contain = FALSE 
)
```
<br>

### Argument <code>neg_values</code>
```{r}
# Check for the absence of negative values. Warning: neg_values = TRUE means "no check". Thus, negative numbers can be present or not. 
arg_check(
    data = mat1, 
    class = "matrix",
    neg_values = FALSE # with TRUE, integers stored as double are accepted.
)
```
<br>

### Argument <code>inf_values</code>
```{r}
# Check for the absence of Inf/-Inf values. Warning: inf_values = TRUE means "no check". Thus, infinite numbers can be present or not. 
arg_check(
    data = mat1, 
    class = "matrix",
    inf_values = FALSE
)
```
<br>

### Argument <code>print</code>
```{r}
# No error message printed because print is FALSE.
res <- arg_check(
    data = mat1, 
    class = "data.frame",
    print = FALSE
)

# Error message printed.
res <- arg_check(
    data = mat1, 
    class = "data.frame",
    print = TRUE
)

# Even if print is TRUE, no error message printed because no error.
res <- arg_check(
    data = mat1, 
    class = "matrix",
    print = TRUE
)
```
<br>

### Arguments <code>data_name</code>
```{r}
# Example
tempo <- arg_check(
    data = vec1, 
    class = "integer", 
    data_name = "OBSERVATION_1"
)
tempo$text

# In fact, these two arguments are interesting when check() is used inside functions
fun1 <- function(arg1){
    tempo <- arg_check(
                data = arg1, 
                class = "integer", 
                data_name = NULL # if NULL, the name displayed is arg1
    )
    if(tempo$problem == TRUE){
        cat(paste0("\n\n================\n\n", tempo$text, "\n\n================\n\n"))
    }
}
fun1(arg1 = vec4) # error message because arg1 requires a vector of integers
```
<br>

### Argument <code>safer_check</code>
```{r}
# Safer checkings performed before main code running.
arg_check(
    data = mat1, 
    class = "matrix", 
    safer_check = TRUE
)
```
<br>

### Argument <code>lib_path</code> and <code>error_text</code>
```{r}
arg_check(
    data = mat1, 
    class = "matrix", 
    lib_path = ".", 
    error_text = " === TEXT ADDED ===" 
)
```
<br>

### All the arguments
```{r}
# See the examples of fun_info() to test different classes of objects
arg_check(
    data = vec1, 
    class = "integer", 
    typeof = NULL, 
    mode = NULL, 
    length = NULL, 
    prop = FALSE, 
    double_as_integer_allowed = FALSE, 
    options = NULL, 
    all_options_in_data = FALSE, 
    na_contain = FALSE, 
    neg_values = TRUE, 
    inf_values = TRUE, 
    print = FALSE, 
    data_name = NULL, 
    data_arg = TRUE, 
    safer_check = TRUE, 
    lib_path = NULL,
    error_text = "" 
)
```
