---
title: "arg_check(): more examples"
---
<br>

```{r, include = FALSE}
knitr::opts_chunk$set(
    error = TRUE, 
    collapse = FALSE,
    comment = "##"
)
```

```{r setup}
library(saferDev)
```
<br>


### Dataset

```{r}
vec1 <- -1:3 # vector of integers.
vec2 <- 1:3 / 3 # vector of positive proportions.
vec3 <- c(1, 2, 3) # vector of positive integers but stored as "double".
vec4 <- "pearson" # vector of characters.
vec5 <- c("a", "b","a", "b") # vector of characters.
mat1 <- matrix(vec1) # matrix of integers.
mat2 <- matrix(c(1:3 / 3, NA)) # matrix of proportions with NA.
```
<br>

### Dataset info

```{r}
vec1 # vector of integers.
vec2 # vector of proportions.
vec3 # vector of integers but stored as "double".
vec4 # vector of characters.
vec5 # vector of characters.
mat1 # matrix of integers.
mat2 # matrix of proportions with NA.
```
<br>

### Simple examples
```{r}
# Check that vec1 is class integer.
arg_check(data = vec1, class = "integer")


# Check that vec1 is a numeric vector -> error because vec1 is not class "numeric", 
# but class "integer".
arg_check(data = vec1, class = "numeric")


# Check that vec1 is an integer vector of length 3 without negative values and 
# without NA -> error because of length 5 and negative values inside vec1.
arg_check(
    data = vec1, 
    class = "vector", 
    typeof = "integer", 
    length = 3, 
    neg_values = FALSE, 
    na_contain = FALSE
)


# No result displayed because the output list is assigned into res
# (see below the print argument).
res <- arg_check(data = vec1, class = "integer")


# With data = NULL, the function systematically report a checking problem.
arg_check(data = NULL, class = "integer")
```
<br>

### Arguments <code>class</code>, <code>typeof</code>, <code>mode</code> and <code>length</code>
These arguments accept as values the classical kind of class, type, mode and length, respectively, except the <code>class</code> argument, which:

- also accepts <code>"vector"</code>, <code>"ggplot2"</code> (i.e., objects of class <code>c("gg", "ggplot")</code> for R < 4.5 and <code>c("ggplot2::ggplot", "ggplot", "ggplot2::gg", "S7_object", "gg")</code> otherwise) or <code>"ggplot_built"</code> (i.e., objects of class <code>"ggplot_built"</code> for R < 4.5 and <code>c("ggplot2::ggplot_built", "ggplot_built", "ggplot2::gg", "S7_object")</code> otherwise) .
- accepts <code>"matrix"</code> for matrices (objects of class <code>c("matrix" "array")</code>).


```{r, echo = TRUE, error = TRUE}
# Example.
arg_check(
    data = vec1, 
    class = "vector", 
    typeof = "integer", 
    mode = "numeric", 
    length = 5, 
)

# Warning: the function does not check for inconsistencies between arguments.
# It just checks if everything is ok between arguments values and the tested object.
# Here the mode "character" exists, is inconsistant with typeof "integer", but is not signaled.
arg_check(
    data = vec1, 
    typeof = "integer", 
    mode = "character"
)

# Bad mode argument value.
arg_check(
    data = vec1, 
    mode = "integer",
)
```
<br>

### Argument <code>prop</code>
```{r}
# Check for values between 0 and 1 only.
arg_check(
    data = vec2, 
    prop = TRUE 
)
```
<br>

### Argument <code>double_as_integer_allowed</code>
```{r}
# Check for integers, even if numbers are stored as double.
# With double_as_integer_allowed = TRUE # with TRUE, integers stored as double are accepted.
arg_check(
    data = vec3, 
    typeof = "integer",
    double_as_integer_allowed = TRUE
)
```
<br>

### Argument <code>options</code>
```{r}
# Check for the content among values. Here no error detected because vec4 contains "pearson".
arg_check(
    data = vec4, 
    options = c("pearson", "spearman", "kendall")
)
```
<br>

### Argument <code>all_options_in_data</code>
```{r}
# Check for the content among values.
# Here no error detected because vec5 indeed contains at least one copy of "a" and "b" and nothing else.
arg_check(
    data = vec5,
    options = c("a", "b"), 
    all_options_in_data = TRUE
)

# Here no error detected because vec5 indeed contains at least one copy of "a" and "b", among other values.
arg_check(
    data = vec5,
    options = c("a", "b", "c"), 
    all_options_in_data = FALSE
)

# Here an error is detected because vec5 must contain "a", "b" and "c", at least once each and nothing else.
arg_check(
    data = vec5,
    options = c("a", "b", "c"), 
    all_options_in_data = TRUE
)
```
<br>

### Argument <code>na_contain</code>
```{r}
# Check for the absence of NA.
# Warning: na_contain = TRUE means "no check". Thus, NA can be present or not. 
arg_check(
    data = mat2, 
    class = "matrix", 
    prop = TRUE,
    na_contain = FALSE 
)
```
<br>

### Argument <code>neg_values</code>
```{r}
# Check for the absence of negative values.
# Warning: neg_values = TRUE means "no check". Thus, negative numbers can be present or not. 
arg_check(
    data = mat1, 
    class = "matrix",
    neg_values = FALSE
)
```
<br>

### Argument <code>inf_values</code>
```{r}
# Check for the absence of Inf/-Inf values.
# Warning: inf_values = TRUE means "no check". Thus, infinite numbers can be present or not.
# Only type "double" can have Inf values (not integers).
arg_check(
    data = mat1, 
    class = "matrix",
    inf_values = FALSE
)
```
<br>

### Argument <code>print</code>
```{r}
# No error message printed because print is FALSE.
res <- arg_check(
    data = mat1, 
    class = "data.frame",
    print = FALSE
)

# Error message printed.
res <- arg_check(
    data = mat1, 
    class = "data.frame",
    print = TRUE
)

# Even if print is TRUE, no error message printed because no error.
res <- arg_check(
    data = mat1, 
    class = "matrix",
    print = TRUE
)
```
<br>

### Argument <code>data_name</code>
```{r}
# Modification of the name of the tested object in the message
arg_check(
    data = vec1, 
    class = "integer", 
    data_name = "OBSERVATION_1"
)

# This argument is interesting when data is not an object.
x <- 1
arg_check(
    data = get("x"), 
    class = "integer", 
    data_name = "x"
)

# If "data_name" is not used.
arg_check(
    data = get("x"), 
    class = "integer", 
    data_name = NULL
)
```
<br>

### Arguments <code>data_arg</code>
```{r}
# Message with "argument". Useful when arg_check() is used inside a function to test its arguments.
arg_check(
    data = vec2, 
    class = "integer", 
    data_arg = TRUE
)

# Message with "object".
arg_check(
    data = vec2, 
    class = "integer", 
    data_arg = FALSE
)
```
<br>

### Argument <code>safer_check</code>
```{r}
# Safer checkings performed before main code running.
arg_check(
    data = mat1, 
    class = "matrix", 
    safer_check = TRUE
)
```
<br>

### Argument <code>lib_path</code>
```{r}
# Error returned because the R required packages are not in the "." path.
```{r}
arg_check(
    data = mat1, 
    class = "matrix", 
    lib_path = "."
)
```
<br>

### Argument <code>error_text</code>
```{r}
# Add a text in error messages returned by all_args_here().
arg_check(
    data = mat1, 
    class = "matrix", 
    lib_path = "."
    error_text = " === TEXT ADDED ==="
)
```
<br>

### All the arguments
```{r}
arg_check(
    data = vec1, # object to check.
    class = "integer", # check the expected class.
    typeof = NULL, # check the expected type.
    mode = NULL,  # check the expected mode.
    length = NULL,  # check the expected length.
    prop = FALSE,  # check the expected proportion.
    double_as_integer_allowed = FALSE,  # if set to TRUE, typeof = "double" or class = "numeric" would not return an error if the object contains integers recorded as double.
    options = NULL,  # check all the possible option values.
    all_options_in_data = FALSE,  # sf TRUE, all of the options of the options argument must be present at least once in the data argument, and nothing else. If FALSE, some or all of the options must be present in the data argument, and nothing else.
    na_contain = FALSE,  # can the data object contain NA (only tested if na_contain = FALSE)?
    neg_values = TRUE,  # can the data object contain negative values (only tested if na_contain = FALSE)?
    inf_values = TRUE,  # can the data object contain infinite values (only tested if na_contain = FALSE)?
    print = FALSE,  # print a message if the $problem output is TRUE? 
    data_name = NULL,  # name of the object to test.
    data_arg = TRUE,  # is the tested object a function argument? If TRUE (default), "ARGUMENT" is written in output messages, otherwise "OBJECT".
    safer_check = TRUE, # perform some "safer" checks (see https://github.com/safer-r)?
    lib_path = NULL, # absolute pathways of the directories containing the required packages if not in the default directories.
    error_text = "" # add information in error messages returned by the function.
)
```
